Investigating Biases Associated with Dietary Starch Incorporation and Retention with an Oral Biofilm Model

### 2022-05-19 analysis/paper/matmet.Rmd
fix citekeys and keep md

Statistical analysis was conducted in `r version$version.string`  [@Rbase] and
the following packages: tidyverse [@tidyverse2019], broom [@R-broom], 
here [@R-here], and patchwork [@R-patchwork].
<!-- R packages used -->

>>>>

Statistical analysis was conducted in `r version$version.string`  [@Rbase] and
the following packages: tidyverse [@tidyverse2019], broom [@Rbroom], 
here [@Rhere], and patchwork [@Rpatchwork].
<!-- R packages used -->

### 2022-05-19 analysis/paper/paper.Rmd
fix citekeys and keep md

This preliminary study shows that a very small proportion of the input starch
granules are retained in a dental calculus model. This and previous studies 
have shown that calculus has a low capacity for retention of starch granules, 
an effect that is compounded by diagenetic effects in archaeological remains, 
resulting in low overall counts of extracted granules. 
The proportion of starches consumed will in many cases be reflected 
in the quantity
of starches extracted from the dental calculus---i.e., the more starch granules
entering the oral cavity, the more will be recovered from extraction---at 
least in modern calculus samples unaffected by diagenesis and hydrolysis.
Whether or not this also applies to archaeological samples remains to be tested.
Additionally, we have
shown that the size of granules will influence the likelihood of incorporation, 
as large (>20 $\mu$m) starches have a decreased incorporation rate, medium 
(10--20 $\mu$m)
starches an increased rate, and small (<10 $\mu$m) granules remained somewhat
constant. The size of calculus deposit also seems to influence the capacity of 
granule incorporation; as the size of the deposit increases, so does the
absolute count of incorporated granules.  
While we have shown multiple factors that influence the likelihood 
of incorporation, the process still appears to be somewhat stochastic. Further
research is needed to make sense of the contributing factors, and to explore the
mechanisms of intra-oral starch incorporation and retention in dental calculus.
The oral biofilm model described in this study provides a method
to explore the incorporation and extraction of dietary compounds from dental calculus 
in a controlled laboratory setting, and unearth the potential biases associated 
with dietary research conducted on archaeological dental calculus.

@@

All scripts and data used in the analysis are available on OSF 
(https://osf.io/uc5qy/) and Github (https://github.com/bbartholdy/byoc-starch),
following the format provided by the rrtools package [@R-rrtools].
More detailed protocols are available on OSF (https://osf.io/akevs/)
and protocols.io (https://www.protocols.io/workspaces/byoc).
Additional tables and figures are available in the Supplementary Material
(https://osf.io/ucxsv/).

>>>>

```{r child="conclusion.Rmd"}

```

@@

All scripts and data used in the analysis are available on OSF 
(https://osf.io/uc5qy/) and Github (https://github.com/bbartholdy/byoc-starch),
following the format provided by the rrtools package [@Rrrtools].
More detailed protocols are available on OSF (https://osf.io/akevs/)
and protocols.io (https://www.protocols.io/workspaces/byoc).
Additional tables and figures are available in the Supplementary Material
(https://osf.io/ucxsv/).
